---
title: "Homework 6 - STAT 5362 Statistical Computing"
author:
  - Sen Yang^[<sen.2.yang@uconn.edu>; M.S. student at
    Department of Statistics, University of Connecticut.]
date: "`r format(Sys.time(), '%d %B %Y')`"
documentclass: article
papersize: letter
fontsize: 11pt
bibliography: template.bib
biblio-style: asa
output:
  bookdown::pdf_document2
abstract: |
    This is homework 6 for STAT 5362 - Statistical Computing.
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## some utility functions, see the source code for details
source("utils_template.R")

## specify the packages needed
pkgs <- c("DT", "leaflet", "splines2", "webshot")
need.packages(pkgs)

## external data can be read in by regular functions,
## such as read.table or load

## for latex and html output
isHtml <- knitr::is_html_output()
isLatex <- knitr::is_latex_output()
latex <- ifelse(isLatex, '\\LaTeX\\', 'LaTeX')

## specify global chunk options
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
                      out.width = "90%", fig.align = "center")

```


# Rejection sampling {#sec:5.2.1}

## Show that $g$ is a mixture of Gamma distributions

$g$ is a probability density on $(0,\infty)$, such that 
$$g(x) \propto (2x^{\theta-1} + x^{\theta-1/2})e^{-x}.$$
With a normalizing constant $C$, we have 
$$C\int_0^{\infty}(2x^{\theta-1} + x^{\theta-1/2})e^{-x} dx=1.$$
Therefore,
\begin {align*}
    & \; C\int_0^{\infty}(2x^{\theta-1} + x^{\theta-1/2})e^{-x} dx \\
    & = 2C\int_0^{\infty}2x^{\theta-1}e^{-x}dx + C\int_0^{\infty}x^{\theta+1/2-1}e^{-x}dx \\
    & = 2C\Gamma(\theta) + C\Gamma(\theta + \frac{1}{2}) \\
    & \Longrightarrow C = \frac{1}{2\Gamma(\theta) + \Gamma(\theta + \frac{1}{2})}
\end {align*}

Let us assume the mixture of the Gamma distribution would like
$$g(x)=c_1\Gamma(\alpha_1,\beta_1)+c_2\Gamma(\alpha_2,\beta_2)$$
Obviously, $\beta_1=\beta_2=1$, and we have
\begin{align*}
    g(x) & = c_1\Gamma(\alpha_1,1)+c_2\Gamma(\alpha_2,1) \\
    & = \frac{c_1}{\Gamma(\alpha_1)}x^{\alpha_1-1}e^{-x} +   
    \frac{c_2}{\Gamma(\alpha_2)}x^{\alpha_2-1}e^{-x} \\
    & = C(2x^{\theta-1} + x^{\theta-1/2})e^{-x} \\
    & \Longrightarrow \alpha_1  = \theta \qquad \alpha_2 = \theta + \frac{1}{2} \\
    & \qquad \frac{c_1}{\Gamma(\alpha_1)}  = 2C \qquad \frac{c_2}{\Gamma(\alpha_2)} = C \\
    & \Longrightarrow c_1=\frac{2\Gamma(\theta)}{2\Gamma(\theta)+\Gamma(\theta+\frac{1}{2})}\\
    & \qquad c_2=\frac{\Gamma(\theta+\frac{1}{2})}{2\Gamma(\theta)+\Gamma(\theta+\frac{1}{2})}
\end {align*}
It means $g$ is a mixture of $\Gamma(\theta,1)$ and $\Gamma(\theta+ \frac{1}{2},1)$ with corresponding weight $c_1=\frac{2\Gamma(\theta)}{2\Gamma(\theta)+\Gamma(\theta+\frac{1}{2})}$, $c_2=\frac{\Gamma(\theta+\frac{1}{2})}{2\Gamma(\theta)+\Gamma(\theta+\frac{1}{2})}$, respectively.
\vspace{5mm}

## Sample from g

The algorithm to generate a sample from g:

\begin{algorithm} [H]
\caption{Generate a sample from g}
\begin{algorithmic} 
\STATE Generate a sample $U \thicksim U(0,1)$
\IF{$U < c_1$}
\STATE $X = \Gamma(\theta,1)$
\ELSE
\STATE $X = \Gamma(\theta+\frac{1}{2},1)$
\ENDIF
\STATE return X.
\end{algorithmic}
\end{algorithm}

Then, we realize this algorithm in R.
```{r sample_g, echo=T}
# Sample from g
## function to generate a sample from g
g_sample <- function(n,theta) {
  c1 <- 2*gamma(theta)/(2*gamma(theta)+gamma(theta+0.5))
  uni <- runif(n)
  samples <- data.frame(spl=matrix(0,n,1))
  samples[which(uni<c1),1] <- rgamma(length(which(uni<c1)),shape = theta,scale = 1)
  samples[which(uni>=c1),1] <- rgamma(length(which(uni>=c1)),shape = theta+0.5,scale = 1)
  samples
}
## generate a sample with theta = 2
g_sample1 <- g_sample(10000,2)

## ture density
trued_g <- function (x,theta) {
  c1 <- 2*gamma(theta)/(2*gamma(theta)+gamma(theta+0.5))
  trued_g <- c1*dgamma(x,shape=theta,scale=1)+(1-c1)*dgamma(x,shape=theta+0.5,scale=1)
  trued_g
}
## plot
library(ggplot2)
ggplot(g_sample1) + stat_function(aes(col="true"), fun=trued_g, args = list(theta=2)) +
  geom_density(aes(spl, col='kernel')) + scale_color_manual(values = c("red", "blue")) +
  labs(title = expression(paste("kernel density vs. true density with ", theta, "=2")),
       x = "x", y = "g(x)") + theme(plot.title = element_text(hjust = 0.5))
```

## Sample from f

To reduce the expected number of iterations, $\alpha$ should be small.
\begin {align*}
  \alpha & = \sup_{x>0} \frac{q(x)}{g(x)} \\
  & = \sup_{x>0} \frac{\sqrt{4+x}x^{\theta-1}e^{-x}}
  {C(2x^{\theta-1}+x^{\theta-1/2})e^{-x}} \\
  & = \frac{1}{C} \sup_{x>0} \frac{\sqrt{4+x}}{(2+\sqrt{x})}
\end {align*}

Since $(\frac{\sqrt{4+x}}{2+\sqrt{x}})^{'}=\frac{2\sqrt{x}-4}{2\sqrt{4x+x^2}(2+\sqrt{x})^2}$, this is a convex function that is decreasing over $(0,4)$ and increasing over $(4,\infty)$. Therefore, 
$$sup_{x>0}(\frac{\sqrt{4+x}}{2+\sqrt{x}})=\max\{\lim_{x\rightarrow0}\frac{\sqrt{4+x}}{2+\sqrt{x}},\lim_{x \rightarrow \infty}\frac{\sqrt{4+x}}{2+\sqrt{x}}\} = 1$$
$$\Longrightarrow \alpha = \sup_{x>0} \frac{q(x)}{g(x)}= \frac{1}{C}$$

The algorithm to generate a sample from f:

\begin{algorithm} [H]
\caption{Generate a sample from f}
\begin{algorithmic} 
\STATE Step 1: Generate a sample from $X \thicksim g$ and $U \thicksim U(0,1)$
\STATE Step 2:
\IF{$U > \frac{q(x)}{\alpha g(x)}$}
\STATE Go to step 1.
\ELSE
\STATE Return X.
\ENDIF
\end{algorithmic}
\end{algorithm}

Then, we realize this algorithm in R.
```{r sample_f, echo=T}
# sample from f
## generate a sample from f
f_sample <- function(n,theta) {
  g_sample <- g_sample(n,theta)
  u_sample <- runif(n)
  c <- 1/(2*gamma(theta)+gamma(theta+0.5))
  f_sample <- matrix(0,n,1)
  for (i in 1:n) {
    qx <- sqrt(4 + g_sample[i,1])*g_sample[i,1]^(theta-1)*exp(-g_sample[i,1])
    gx <- (2*g_sample[i,1]^(theta-1)+g_sample[i,1]^(theta-0.5))*exp(-g_sample[i,1])
    if (u_sample[i] > qx*c/gx) next
      else {f_sample[i,1]<-g_sample[i,1]}
  }
  f_sample<-as.data.frame(f_sample[-which(f_sample==0),1])
  colnames(f_sample)<-"f"
  f_sample
}

f_sample1<-f_sample(10000,2)

qx<- function(x, theta) {
  qx<-sqrt(4 + x)*x^(theta-1)*exp(-x)
  qx
}
trued_f <- function (x,theta) {
  intgrl<-integrate(function(x) qx(x,theta),0,Inf)
  c_prime <-1/intgrl[["value"]]
  trued_f <- c_prime*sqrt(4 + x)*x^(theta-1)*exp(-x)
  trued_f
}

## plot
library(ggplot2)
ggplot() + stat_function(aes(col="true"), fun=trued_f, args = list(theta=2)) +
  geom_density(data=f_sample1,aes(f,col="kernel")) +
  scale_color_manual(values = c("red","blue")) +
  labs(title = expression(paste("kernel density vs. true density with ", theta, "=2")),
       x = "x", y = "f(x)") + theme(plot.title = element_text(hjust = 0.5))
```

# Mixture Proposal {#sec:5.2.2}

## Generate a sample using a mixture of beta distribution as the instrumental density

Since $0<x<1$, $0<x^2<1$. Then, we have 
$$q(x)=\frac{x^{\theta-1}}{1+x^2}+\sqrt{2+x^2}(1-x)^{\beta-1}<x^{\theta-1}+\sqrt{3}(1-x)^{\beta-1}=h(x).$$
Therefore, we choose 
$$h(x)=x^{\theta-1}+\sqrt{3}(1-x)^{\beta-1} \qquad with \quad\alpha=1.$$

where $g(x)=Ch(x)=C(x^{\theta-1}+\sqrt{3}(1-x)^{\beta-1})=p_1\theta x^{\theta-1}+p_2\beta (1-x)^{\beta-1} \thicksim p_1 Beta(\theta,1)+p_2Beta(1,\beta)$.

Then,
\begin {align*}
  g(x) & = \int_0^1 C(x^{\theta-1}+\sqrt{3}(1-x)^{\beta-1})dx \\
  & = \frac{C}{\theta}x^{\theta}-\frac{\sqrt{3}C}{\beta}(1-x)^{\beta} \vert_{x=0}^{1} \\
  & = \frac{C}{\theta}+\frac{\sqrt{3}C}{\beta} \\
  & = 1 \\
  & \Longrightarrow C=\frac{\theta\beta}{\beta+\sqrt{3}\theta} \\
  & \Longrightarrow p_1=\frac{\beta}{\beta+\sqrt{3}\theta} \qquad p_2 = 
  \frac{\sqrt{3}\theta}{\beta+\sqrt{3}\theta}
\end {align*}

The algorithm to generate a sample from f:

\begin{algorithm} [H]
\caption{Generate a sample from f}
\begin{algorithmic} 
\STATE Step 1: Generate a sample from g(x)
\STATE Generate a sample $U \thicksim U(0,1)$
\IF{$U < p_1$}
\STATE $X = Beta(\theta,1)$
\ELSE
\STATE $X = Beta(1,\beta)$
\ENDIF
\STATE return X.
\STATE Step 2: Generate a sample from $X \thicksim g$ and $U \thicksim U(0,1)$
\STATE Step 3:
\IF{$U > \frac{q(x)}{\alpha h(x)}=\frac{q(x)}{ h(x)}$}
\STATE Go to step 2.
\ELSE
\STATE Return X.
\ENDIF
\end{algorithmic}
\end{algorithm}

Then, we realize this algorithm in R.
```{r sample_f_5.2.2, echo=T}
# Sample from g
## function to generate a sample from g
g_sample <- function(n,theta,beta) {
  p1 <- beta/(beta+sqrt(3)*theta)
  uni <- runif(n)
  samples <- data.frame(spl=matrix(0,n,1))
  samples[which(uni<p1),1] <- rbeta(length(which(uni<p1)),shape1 = theta, shape2 = 1)
  samples[which(uni>=p1),1] <- rbeta(length(which(uni>=p1)),shape1 = 1, shape2 = beta)
  samples
}

# sample from f
## generate a sample from f
f_sample <- function(n,theta,beta) {
  g_sample <- g_sample(n,theta,beta)
  u_sample <- runif(n)
  f_sample <- matrix(0,n,1)
  for (i in 1:n) {
    qx <- g_sample[i,1]^(theta-1)/(1+g_sample[i,1]^2)+
      sqrt(2+g_sample[i,1]^2)*(1-g_sample[i,1])^(beta-1)
    hx <- g_sample[i,1]^(theta-1)+3^0.5*(1-g_sample[i,1])^(beta-1)
    if (u_sample[i] > qx/hx) next
    else {f_sample[i,1]<-g_sample[i,1]}
  }
  f_sample<-as.data.frame(f_sample[-which(f_sample==0),1])
  colnames(f_sample)<-"f"
  f_sample
}

f_sample1<-f_sample(10000,2,4)

qx<- function(x, theta, beta) {
  qx<-x^(theta-1)/(1+x^2)+sqrt(2+x^2)*(1-x)^(beta-1)
  qx
}
trued_f <- function (x,theta,beta) {
  intgrl<-integrate(function(x) qx(x,theta,beta),0,1)
  c_prime <-1/intgrl[["value"]]
  trued_f <- c_prime*(x^(theta-1)/(1+x^2)+sqrt(2+x^2)*(1-x)^(beta-1))
  trued_f
}

## plot
library(ggplot2)
ggplot() + stat_function(aes(col="true"), fun=trued_f, args = list(theta=2, beta=4)) +
  geom_density(data=f_sample1,aes(f,col="kernel")) +
  scale_color_manual(values = c("red","blue")) +
  labs(title = expression(paste("kernel density vs. true density with ", theta, 
                                "=2 and ", beta, "=4")),
       x = "x", y = "f(x)") + theme(plot.title = element_text(hjust = 0.5))
```

## Generate a sample using individual Beta distributions

Let
$$q_1(x)=\frac{x^{\theta-1}}{1+x^2} \qquad q_2(x)=\sqrt{2+x^2}(1-x)^{\beta-1}$$
$$h_1(x)=x^{\theta-1} \qquad h_2(x)=\sqrt{3}(1-x)^{\beta-1} \qquad with\:\: \alpha_1=\alpha_2=1. $$
Therefore, we can get corresponding $g(x)$.
$$g_1(x)=\theta x^{\theta-1} \qquad g_2(x)=\beta (1-x)^{\beta-1}$$

The algorithm to generate a sample from f:

\begin{algorithm} [H]
\caption{Generate a sample from f}
\begin{algorithmic} 
\STATE Step 1: Generate a sample $U \thicksim U(0,1)$
\STATE Step 2:
\IF{$U < \frac{\alpha_1}{\alpha_1+\alpha_2}=0.5$}
\STATE Sample from $X \thicksim g_1$ and $U_1 \thicksim U(0,1)$
\STATE Step 3: 
  \IF{$U>\frac{q_1(x)}{\alpha_1h_1(x)}=\frac{q_1(x)}{h_1(x)}$}
  \STATE Go to Step 3.
  \ELSE
  \STATE Return X.
  \ENDIF
\ELSE
\STATE Sample from $X \thicksim g_2$ and $U_2 \thicksim U(0,1)$
\STATE Step 4: 
  \IF{$U>\frac{q_2(x)}{\alpha_2h_2(x)}=\frac{q_2(x)}{h_2(x)}$}
  \STATE Go to Step 4.
  \ELSE
  \STATE Return X.
  \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

Then, we realize this algorithm in R.
```{r sample_f_5.2.2_b, echo=T}
# Separate Beta

# Sample from g
## function to generate a sample from g
g_sample1 <- function(n,theta,beta) {
  g_sample1 <- data.frame(rbeta(n,shape1=theta,shape2=1))
  g_sample1
}

g_sample2 <- function(n,theta,beta) {
  g_sample2 <- data.frame(rbeta(n,shape1=1,shape2=beta))
  g_sample2
}

# sample from f
## generate a sample from f
f_sample_sep <- function(n,theta,beta) {
  u_sample <- runif(n)
  n1 <- length(which(u_sample<0.5))
  g_sample1 <- g_sample1(n1,theta,beta)
  u_sample1 <- runif(n1)
  f_sample1 <- matrix(0,n1,1)
  for (j in 1:n1) {
    qx1 <- g_sample1[j,1]^(theta-1)/(1+g_sample1[j,1]^2)
    hx1 <- g_sample1[j,1]^(theta-1)
    if (u_sample1[j] > qx1/hx1) next
     else {f_sample1[j,1]<-g_sample1[j,1]}
   }
  n2 <- length(which(u_sample>=0.5))
  g_sample2 <- g_sample2(n2,theta,beta)
  u_sample2 <- runif(n2)
  f_sample2 <- matrix(0,n2,1)
  for (k in 1:n2) {
    hx2 <- 3^0.5*(1-g_sample2[k,1])^(beta-1)
    qx2 <- sqrt(2+g_sample2[k,1]^2)*(1-g_sample2[k,1])^(beta-1)
    if (u_sample2[k] > qx2/hx2) next
    else {f_sample2[k,1]<-g_sample2[k,1]}
    }
  f_sample<-rbind(f_sample1,f_sample2)
  f_sample<-as.data.frame(f_sample[-which(f_sample==0),1])
  colnames(f_sample)<-"f"
  f_sample
}  

f_sample_sep<-f_sample_sep(10000,2,4)

## plot
library(ggplot2)
ggplot() + stat_function(aes(col="true"), fun=trued_f, args = list(theta=2, beta=4)) +
  geom_density(data=f_sample_sep,aes(f,col="kernel")) +
  scale_color_manual(values = c("red","blue")) +
  labs(title = expression(paste("kernel density vs. true density with ", theta, 
                                "=2 and ", beta, "=4")),
       x = "x", y = "f(x)") + theme(plot.title = element_text(hjust = 0.5))
```







